---
title: "Work Distractions"
subtitle: "SAT 231: Calendar Query"
author: "Ziji Zhou"
date: "Last updated \\today"
output: pdf_document
---

```{r setup, include = FALSE}
# load packages
library(tidyverse)
library(lubridate)
library(ical)
library(RColorBrewer)
library(kableExtra)

# set code chunk defaults
knitr::opts_chunk$set(tidy = F, # display code as typed
                      size = "small", # slightly smaller code font
                      message = FALSE,
                      warning = FALSE,
                      comment = "\t") 

# set black & white default plot theme 
# theme_set(theme_classic()) # uncomment this line if desired

# improve digit and NA display 
options(scipen = 1, knitr.kable.NA = '')
```

<!--
Please remove my comments from the code chunks and customize the code chunks on your own, including chunk labels and options and code documentation.

You can leave my markdown comments for guidance in writing your report.

Please rename the headings. 
You may add subheadings as desired to break up the report. 
-->

# How Much Distractions Do I Have During Work <!-- (e.g., How do I spend my time?) -->

<!--
The code below is provided as a starting point for importing an `ics` file into R as a dataframe, doing some initial wrangling, and adding up the time spent on each activity. You are not required to use this code, so delete or modify it as you see fit. 

If you are using the code, be sure to update the capitalized parts with the relevant path to where your `ics` file is located, and set `eval = TRUE`.

As a reminder here's the link for exporting a Google calendar to ics (if on a Mac, you can hold cmd + click to open the page):
https://support.google.com/calendar/answer/37111?hl=en&ref_topic=10509946
-->
```{r, eval = TRUE}
# Data import and preliminary wrangling
distractions_data <- "Distractions.ics" %>% 
  # Use ical package to import into R
  ical_parse_df() %>% 
  # Convert to "tibble" data frame format
  as_tibble() %>% 
  mutate(
    # Use lubridate package to wrangle dates and times
    start_datetime = with_tz(start, tzone = "America/New_York"),
    end_datetime = with_tz(end, tzone = "America/New_York"),
    # duration_sec = end_datetime - start_datetime,
    duration_sec = difftime(end_datetime, start_datetime, units = "secs"),
    # Convert calendar entry to all lowercase and rename
    activity = summary)  %>%
  # Pivot longer into start and end time
  pivot_longer(c("start","end"), names_to = "Start/End", values_to = "time") %>%
  mutate(
    # Get time of day
    time_of_day = hms::hms(seconds = second(time), minutes = minute(time), hours = hour(time)),
    # Get date as a categorical variable
    date = floor_date(time, unit = "day"),
    date_chr = paste(as.character(month(date)), as.character(day(date)), as.character(year(date)), sep = "/"),
    type = NA,
    class = NA,
    work_uid = NA
  )

schoolwork_data <- "Schoolwork.ics" %>% 
  # Use ical package to import into R
  ical_parse_df() %>% 
  # Convert to "tibble" data frame format
  as_tibble() %>% 
  mutate(
    # Use lubridate package to wrangle dates and times
    start_datetime = with_tz(start, tzone = "America/New_York"),
    end_datetime = with_tz(end, tzone = "America/New_York"),
    # duration_sec = end_datetime - start_datetime,
    duration_sec = difftime(end_datetime, start_datetime, units = "secs"),
    # Convert calendar entry to all lowercase and rename
    activity = summary,
    # Split activity column into work and class columns
    type = substring(activity, 0, regexpr(", ", activity) - 1),
    class = substring(activity, regexpr(", ", activity) + 2))  %>%
  # Pivot longer into start and end time
  pivot_longer(c("start","end"), names_to = "Start/End", values_to = "time") %>%
  mutate(
    # Get time of day
    time_of_day = hms::hms(seconds = second(time), minutes = minute(time), hours = hour(time)),
    # Get date as a categorical variable
    date = floor_date(time, unit = "day"),
    date_chr = paste(as.character(month(date)), as.character(day(date)), as.character(year(date)), sep = "/")
  )

# Attach type of work to distractions
for(i in 1:nrow(distractions_data)){
  
  for(x in 1:nrow(schoolwork_data)){
    
    if(x%%2 == 1 && i%%2 == 1){
      if(as.POSIXct(distractions_data$time[i]) %within% interval(as.POSIXct(schoolwork_data$time[x]),as.POSIXct(schoolwork_data$time[x+1]))){
        distractions_data$class[i] = as.character(schoolwork_data$class[x])
        distractions_data$type[i] = as.character(schoolwork_data[x,"type"])
        distractions_data$work_uid[i] = as.character(schoolwork_data[x,"uid"])
        distractions_data$class[i+1] = as.character(schoolwork_data$class[x])
        distractions_data$type[i+1] = as.character(schoolwork_data[x,"type"])
        distractions_data$work_uid[i+1] = as.character(schoolwork_data[x,"uid"])
      }
    }
  }
}

```


```{r, eval = FALSE}
# Display the data
schoolwork_data
distractions_data
activities
distractions

```

<!--
Describe your question(s) here.  Briefly describe your data collection process, including how you defined variables of interest.
-->

* What type of distractions do I interrupt each types of work and for which class? 
* In which class am I the most productive? What type of work? What time of day? (Productiveness is defined by the amount of distractions during that period of work)

  My data collection process involved logging the each time I sit down seriously to do schoolwork. I would take note the class and type of work (split between problems, reading, and writing). For each time I go away from my work I would take note of when I started going off track and when I returned as a seperate data set. My distractions were categorically split into phone, text, talk, food, web, restroom, music.

<!--
Describe what information is conveyed through data visualization #1 (that you'll create below) here.
-->

This visualization presents my work over the two weeks of data collection. The x axis is the date and the y axis indicates the time of that day. Each type of work is divided up by color and class faceted into my four rows to help present which class am I the most productive (or as the data came out to be, unproductive). The orange streaks within the wider transparent blocks represent each distraction and the proportion of the work time it took up.


```{r, fig.width=15, fig.height=20}
# Code for data visualization #1
schoolwork_data %>%
  ggplot(aes(x = as.POSIXct(time_of_day), y = reorder(date_chr,desc(date_chr)))) +
  scale_x_datetime(date_breaks = "1 hours",
                   date_labels = "%H:%M",
                   expand = c(0, 0)) +
  geom_line(aes(color = type, 
                group = uid), 
            size = 10, 
            alpha = 0.3) + 
  geom_line(data = distractions_data ,
            aes(group = uid,
                color = "Distractions"), 
            size = 3, 
            alpha = 0.8) +
  facet_wrap(~class, nrow = 4) +
  labs(title = "Timeline of my Schoolwork and Distractions", 
       x = "Hour",
       y = "Date",
       color = "Colors")
  
# Be sure to provide meaningful title and axes labels 
```

<!--
Describe what information is conveyed through data visualization #2 (that you'll create below) here.
-->

This faceted bar graph shows the distribution of my time with each class and my distractions during them. The y axis displays the minutes while the x is the class the bar represents. The graph is faceted for each type of work and the corresponding distractions. Each distraction is then divided up into color to show which type of distraction is most common for each type of work.

```{r,fig.width=15, fig.height=8}
schoolwork_summary <- schoolwork_data %>%
  group_by(type, class) %>%
  summarize(duration_sec = sum(duration_sec)/2) %>%
  arrange(desc(duration_sec))

distractions_summary <- distractions_data %>%
  group_by(type, class, summary) %>%
  summarize(duration_sec = sum(duration_sec)/2) %>%
  arrange(desc(duration_sec))

# Code for data visualization #2
schoolwork_summary %>%
  ggplot(aes(x = type, y = duration_sec/60)) +
  geom_col(width = 0.9, alpha = 0.5, fill = "red") +
  geom_col(data = distractions_summary, aes(fill = summary), width = 0.75) +
  scale_fill_brewer(palette = "Set2") +
  facet_wrap(~class, nrow = 1) +
  scale_fill_discrete(name = "Distractions") +
  labs(title = "Productivity of each class",
       x = "Class",
       y = "Time spent in minutes",
       color = "Distractions")
  
# Be sure to provide meaningful title and axes labels 
```

<!--
Describe what information is conveyed through the table (that you'll create below) here.

For nicer tables, check out:

* the kable function in the knitr package + the kableExtra package  
  https://bookdown.org/yihui/rmarkdown-cookbook/kable.html 
  https://bookdown.org/yihui/rmarkdown-cookbook/kableextra.html

or 

* the xtable package
  https://cran.r-project.org/web/packages/xtable/vignettes/xtableGallery.pdf
-->

My table is a summary of every individual work session and its corresponding distractions, seperated by class. The table shows the total sessions and distractions, average time per distraction and work session, and finally the proportion of time I spent each session distracted.

```{r, results = 'asis'}
schoolwork_total <- schoolwork_data %>%
  group_by(class) %>%
  summarise(total_time_spent = sum(duration_sec),
            number = n()/2,
            average_work = mean(duration_sec/2))

distractions_total <- distractions_data %>%
  group_by(class) %>%
  summarise(total_time_distracted = sum(duration_sec), 
            number = n()/2,
            average_distraction = mean(duration_sec/2))

total_summary <- schoolwork_total%>%
  inner_join(distractions_total, by = c("class" = "class")) %>%
  mutate(proportion_distracted = as.double(total_time_distracted)/as.double(total_time_spent)) %>%
  arrange(desc(total_time_spent))

total_summary %>%
  kable(col.names = c("Class", "Total time spent", "Total times worked", "Average time per session", "Total time distracted", "Total distractions", "Average time per distraction", "Proportion distracted"), digits = 2, booktabs = TRUE)
# Code for table
```

<!--
To conclude, briefly summarize what you found in response to the questions posed here.
-->

Through my data collection and visualization I was able to get an idea of what leads to the most distractions. Surprisingly my reading for art market had the least amount of distractions, an explanation could be because it has the least work and every time I start reading I have an end in sight. It seems that the afternoon period is the least productive for me with lots of distractions during that period. In addition, using my phone is the most common and time consuming distraction, showing up no matter time period nor class. I think by silencing my phone and placing it far away could do wonders for my productivity. Lastly, short periods of study contains the least time spent being distracted, perhaps short burst sessions of work is better for me?

\newpage
# YOUR REFLECTION HEADING <!-- e.g., Reflection -->

<!--
Write your one-page reflection here in paragraph form.  In particular, address:

* What difficulties in the data collection and analysis process did you encounter?  Identify two of your main hurdles in gathering accurate data.

* What implications does that have for future data collection and/or analysis projects? 

* How much data do you think you'd need to collect in order to answer your question(s) of interest? Would it be hard to collect that data? Why or why not?

* As someone who provides data, what expectations do you have when you give your data (e.g. to Facebook, Google, MapMyRun, etc.)?  

* As someone who analyzes others' data, what ethical responsibilities do you have?
-->



